---
title: "Hands-on_Ex04_2"
author: "Huang Zihan"
date: "May 7, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
---

# 4-2 Visual Statistical Analysis

## 1. Installing and Loading Packages

ggstatsplot is an extension of ggplot2 package for creating graphics with details from statistical tests included in the information-rich plots themselves.

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## 2. Importing Data

```{r}
exam <- read_csv("Exam_data.csv")
```

## 3. One-sample test: gghistostats() method

```{r}
set.seed(1234)

gghistostats(data = exam, x = ENGLISH, type = "bayes",
             test.value = 60, xlab = "English scores")
```

## 4. Two-sample mean test: ggbetweenstats()

```{r}
ggbetweenstats(data = exam, x = GENDER, y = MATHS,
               type = "np", messages = FALSE)
```

## 5. Oneway ANOVA Test: ggbetweenstats() method

```{r}
ggbetweenstats(data = exam, x = RACE, y = ENGLISH,
               type = "p", mean.ci = TRUE,
               pairwise.comparisons = TRUE,
               pairwise.display = "s",
               p.adjust.method = "fdr",
               messages = FALSE)
```

## 6. Significant Test of Correlation: ggscatterstats()

```{r}
ggscatterstats(data = exam, x = MATHS, y = ENGLISH,
               marginal = FALSE)
```

## 7. Significant Test of Association (Depedence) : ggbarstats() methods

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins =
           cut(MATHS, breaks = c(0,60,75,85,100)))
```

```{r}
ggbarstats(exam1, x = MATHS_bins, y = GENDER)
```
