---
title: "Hands-on Exercise 1"
author: "Huang Zihan"
---

# Getting Started

## Install and Launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the laptop. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

# Essential Grammatical Elements in ggplot2: Data

```{r}
ggplot(data = exam_data)
```

# Essential Grammatical Elements in ggplot2: Aesthetic mappings

```{r}
ggplot(data = exam_data,
       aes(x = MATHS))
```

# Essential Grammatical Elements in ggplot2: geom

## Geometric Objects: geom_bar

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_bar()
```

## Geometric Objects: geom_dotplot

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_dotplot(binwidth = 2.5,
               dotsize = 0.5) +
  scale_y_continuous(NULL,
                     breaks = NULL)
```

## Geometric Objects: geom_histogram

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 color = "white",
                 fill = "light green")
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           fill = RACE)) +
  geom_histogram(bins = 20,
                 color = "grey30")
```

## Geometric Objects: geom-density

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_density()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           color = RACE)) +
  geom_density()
```

## Geometric Objects: geom_boxplot

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_boxplot()
```

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_boxplot(notch = TRUE)
```

## Geometric Objects: geom_violin

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_violin()
```

## Geometric Objects: geom_point

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point()
```

## Geom Objects can be combined

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_boxplot() +
  geom_point(position = "jitter",
             size = 1)
```

# Essential Grammatical Elements in ggplot2: stat

## Adding mean value by using stat_summary() and overriding the default geom

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_boxplot() +
  stat_summary(geom = "point",
               fun = "mean",
               color = "light green",
               size = 3)
```

## Adding mean value by using geom\_() and overriding the default stat

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = RACE)) +
  geom_boxplot() +
  geom_point(stat = "summary",
             fun = "mean",
             color = "light pink",
             size = 3)
```

## Best fit curve: geom_smooth()

### Default method used is loess

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(linewidth = 0.5)
```

### The default smoothing method can be overridden

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              linewidth = 0.5)
```

# Essential Grammatical Elements in ggplot2: Facets

## facet_wrap()

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20) +
    facet_wrap(~ RACE)
```

## facet_grid()

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20) +
    facet_grid(~ RACE)
```

# Essential Grammatical Elements in ggplot2: Coordinates

## coord_flip(): to change the bar chart from vertical form to horizontal form

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip()
```

## coord_cartesian(): to fix the axis range

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              linewidth = 0.5) +
  coord_cartesian(xlim = c (0,100),
                  ylim = c (0,100))
```

# Essential Grammatical Elements in ggplot2: Themes

theme() can be picked from the existing themes list, the followings are some examples.

## theme_light()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_light()
```

## theme_classic()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

## theme_gray()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

## theme_minimal()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
