---
title: "Hands-on_Ex06"
author: "Huang Zihan"
date: "May 21, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
---

# Visualizing and Analyzing Time-oriented Data

## 1. Loading R Packages

```{r}
pacman:: p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table, tidyverse, CGPfunctions, ggHoriPlot)
```

## 2. Plotting Calendar Heatmap

### 2.1 Importing Data

```{r}
attacks <- read_csv("eventlog.csv")
```

### 2.2 Examining the data structure

`kable()` can be used to review the structure of the imported data frame

In the following data structure:

|   | Stored Value | Data Format |
|:----------------------:|:----------------------:|:----------------------:|
| timestamp | date-time | POSIXct format |
| source_country | the source of the attack | ISO 3166-1 alpha-2 country code |
| tz | time zone of the source IP address |  |

```{r}
kable(head(attacks))
```

### 2.3 Data Preparation

#### Step 1: Deriving weekday and hour of day fields

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

#### Step 2: Deriving the attacks tibble data frame

`mutate()` of dplyr package is used to convert weekday and hour fields into factor so they’ll be ordered when plotting

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

Table below shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

### 2.4 Building the Calendar Heatmaps

-   `na.omit()` is used to exclude missing value.

-   `geom_tile()` is used to plot tiles (grids) at each x and y position.

-   `theme_tufte()` of ggthemes package is used to remove unnecessary chart junk.

-   `coord_equal()` is used to ensure the plot will have an aspect ratio of 1:1.

-   `scale_fill_gradient()` function is used to creates a two colour gradient (low-high).

```{r}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

### 2.5 Building Multiple Calendar Heatmaps

#### Step 1: Deriving attack by country object

```{r}
attacks_by_country <- count(
  attacks, source_country) %>% # count the number of attacks by country
  mutate(percent = percent(n/sum(n))) %>%
  # calculate the percent of attackes by country
  arrange(desc(n)) # save the results in a tibble data frame
```

#### Step 2: Preparing the tidy data frame

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

#### Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package

```{r}
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## 3. Plotting Cycle Plot

### 3.1 Importing Data

```{r}
air <- read_excel("arrivals_by_air.xlsx")
```

### 3.2 Deriving month and year fields

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### 3.3 Extracting the target country

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### 3.4 Computing year average arrivals by month

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### 3.5 Plotting the cycle plot

```{r}
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
```

## 4. Plotting Slopegraph

### 4.1 Importing Data

```{r}
rice <- read_csv("rice.csv")
```

### 4.2 Plotting the slopegraph

```{r}
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Huang Zihan")
```
