---
title: "Hands-on_Ex08-3"
format: html
author: "Huang Zihan"
date: "Jun 9, 2025"
date-modified: "last-modified"
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
---

# Analytical Mapping

## Installing and loading packages

```{r}
pacman::p_load(tmap, tidyverse, sf)
```

# Importing Data

For this exercise, we use `NGA_wp.rds`, a polygon dataset of Nigeria’s water points at the LGA level, located in the `rds` subdirectory.

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

# Basic Choropleth Mapping

## Visualising distribution of non-functional water point

Plot a choropleth map showing the distribution of non-function water point by LGA

```{r}
p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs")
```

```{r}
p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs")
```

```{r}
tmap_arrange(p2, p1, nrow = 1)
```


# Choropleth Map for Rates

Mapping rates instead of counts is crucial because water points are unevenly distributed, and using counts alone may misrepresent the actual spatial patterns of interest.

## Deriving Proportion of Functional Water Points and Non-Functional Water Points

We use `mutate()` from the dplyr package to calculate `pct_functional` and `pct_nonfunctional`, representing the proportions of functional and non-functional water points per LGA.

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

##  Plotting map of rate

Plot a choropleth map showing the distribution of percentage functional water point by LGA

```{r}
tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs")
```


# Extreme Value Maps

Extreme value maps are a type of choropleth map that emphasize outliers by highlighting the lowest and highest values, aligning with spatial exploratory data analysis (Anselin, 1994).

##  Percentile Map

A percentile map is a specialized quantile map with six categories, using breakpoints from `quantile()` with cumulative probabilities `c(0, .01, .1, .5, .9, .99, 1)`, ensuring both endpoints are included.

### Data Preparation

Step 1: Exclude records with NA by using the code chunk below.

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

### Creating the get.var function

We define an R function to extract a variable (e.g., wp_nonfunctional) as a named vector from an sf data frame using vname (as a string) and df as inputs.

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

### A percentile mapping function

Next, we will write a percentile mapping function by using the code chunk below.

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```


### Test drive the percentile mapping function

To run the function, type the code chunk as shown below.

```{r}
percentmap("total_wp", NGA_wp)
```

## Box map

A box map extends a quartile map by adding lower and upper outlier categories, adjusting breakpoints based on whether outliers fall below or above the calculated fences.

```{r}
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```


### Creating the boxbreaks function

The following R function creates 7 breakpoints for a box map by computing quartiles and fences from a vector `v`, using a multiplier `mult` (default 1.5) for the IQR.

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```


### Creating the get.var function

The get.var function extracts a variable (given as a string) from an sf data frame and returns it as an unnamed vector.

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```


### Test drive the newly created function
Let’s test the newly created function

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```


### Boxmap function

The following R function creates a box map from an sf polygon layer using a specified variable, with customizable legend title, map title, and IQR multiplier, and returns a tmap object.

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
```

```{r}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```

