---
title: "Hands-on_Ex02"
author: "Huang Zihan"
---

# Beyond ggplot2 Fundamentals

## 1. Data Preparation

### Installing and loading the required libraries

```{r}
pacman::p_load(ggrepel,
               ggthemes,
               hrbrthemes,
               patchwork,
               tidyverse)
```

### Importing Data

```{r}
exam_data<-read_csv("Exam_data.csv")
```

## 2. ggrepel: to repel overlapping text

### without ggrepel

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point()+
  geom_smooth(method = lm,
              linewidth=0.5)+
  geom_label(aes(label=ID),
             hjust=.5,
             vjust=-.5)+
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))+
  ggtitle("English scores versus Maths scores for primary 3")
```

### with ggrepel

#### geom_label_repel: add a text box background
```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point()+
  geom_smooth(method = lm,
              linewidth=0.5)+
  geom_label_repel(aes(label = ID),
                   fontface = "bold")+
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))+
  ggtitle("English scores versus Maths scores for primary 3")
```

#### geom_text_repel: display plain text
```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH))+
  geom_point()+
  geom_smooth(method = lm,
              linewidth=0.5)+
  geom_text_repel(aes(label = ID),
                   fontface = "bold",
                  size = 3.5,
                  color = "blue")+
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))+
  ggtitle("English scores versus Maths scores for primary 3")
```

## 3. ggthemes

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = "white",
                 fill = "grey")+
  ggtitle("Distribution of Maths scores")+
  theme_wsj()+
  theme(plot.title = element_text(size = 20))
```

## 4. hrbrthemes

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = "white",
                 fill = "grey")+
  ggtitle("Distribution of Maths scores")+
  theme_ipsum(axis_title_size = 15,
              base_size = 15,
              grid = "N")
```
## 5. pathwork: to combine single graphs

### Create single graphs

```{r}
p1<-ggplot(data = exam_data,
           aes(x=MATHS))+
  geom_histogram(bins=20,
                 boundary=100,
                 color="grey25",
                 fill="grey90")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of Maths Scores")
p1
```

```{r}
p2<-ggplot(data = exam_data,
           aes(x=ENGLISH))+
  geom_histogram(bins=20,
                 boundary=100,
                 color="grey25",
                 fill="grey90")+
  coord_cartesian(xlim=c(0,100))+
  ggtitle("Distribution of English Scores")
p2
```

```{r}
p3<-ggplot(data = exam_data,
           aes(x=MATHS,
               y=ENGLISH))+
  geom_point()+
  geom_smooth(method=lm,
              size=0.5)+
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))+
  ggtitle("English scores versus Maths scores for Primary 3")
p3
```

### Combine MATHS histogram & ENGLISH histogram

```{r}
p1+p2
```

### Combine 3 graphs

1. “/” operator to stack two ggplot2 graphs
2. “|” operator to place the plots beside each other
3. “()” operator the define the sequence of the plotting

```{r}
(p1 / p2) | p3
```

#### plot_annotation(): auto-tagging
```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

#### inset_element(): insert p2 to p3
```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

#### patchwork & ggtheme
```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_wsj()+theme(plot.title = element_text(size = 6))
```